---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# slrmodel

<!-- badges: start -->
[![R-CMD-check](https://github.com/lyvqian/slrmodel/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lyvqian/slrmodel/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/lyvqian/slrmodel/branch/main/graph/badge.svg)](https://app.codecov.io/gh/lyvqian/slrmodel?branch=main)
<!-- badges: end -->

## Overview

The simple linear regression model $y={\beta_0}+{\beta_1}x+{\epsilon}$ is a model that summarizes the relationship between an independent variable `x` (predictor) a dependent variable `y` (response). ${\beta_0}$ and ${\beta_1}$ are usually unknown and need to be estimated using observed data (x~1~,y~1~), (x~2~,y~2~),..., (x~n~,y~n~), and ${\epsilon}$ is a random error term. Deriving the estimators ${\hat{\beta_0}}$ and ${\hat{\beta_1}}$ by the method of least squares would help to obtain the best-fitting line ${\hat{y}}={\hat{\beta_0}}+{\hat{\beta_1}}x$, that is, the straight line demonstrating the best approximation of the given set of data. 

The goal of slrmodel is to calculate the ${\hat{\beta_0}}$ and ${\hat{\beta_1}}$ using the method of least squares, test the significance of (x,y) association, and evaluate the overall fitness of the ${\hat{y}}={\hat{\beta_0}}+{\hat{\beta_1}}x$ model. 

* `slr_beta()` creates a matrix of coefficients of the best fitting line. 
* `slr_rsquared()` calculates the coefficient of determination R^2^.

## Installation

You can install the development version of slrmodel from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lyvqian/slrmodel")
```

## Usage

This is a basic example which shows you how to use the slrmodel package:

```{r}
library(slrmodel)

slr_beta(iris$Petal.Length,iris$Petal.Width) #slr_beta(x, y)

slr_rsquared(iris$Petal.Length,iris$Petal.Width) #slr_rsquared(x, y)
```
Both `slr_beta()` and `slr_rsquared()` have two arguments, `x` and `y`. `x` is a vector containing all observed predictors, while `y` is a vector containing all corresponding responses.

The result of `slr_beta()` is a matrix with 2 rows and 4 columns. The upper row provides ${\hat{\beta_0}}$, the standard error of ${\beta_0}$, the t statistic for testing whether ${\hat{\beta_0}}$ significantly differs from 0, and the p value for determining the significance. The lower row provides the same coefficients for ${\hat{\beta_1}}$.

The result of `slr_rsquared()` is R^2^, the coefficient of determination. It tells the proportion of variation in `y` that is explained by the ${\hat{y}}={\hat{\beta_0}}+{\hat{\beta_1}}x$ model.

With ${\hat{\beta_0}}$ and ${\hat{\beta_1}}$, we can plot the best fitting line, which is ${\hat{y}}=-0.363+0.416x$ in this example: 

```{r pressure, echo = FALSE}
plot(iris$Petal.Length,iris$Petal.Width,xlab="The Petal lengths of iris flowers",ylab="The Petal widths of iris flowers")
intercept = slr_beta(iris$Petal.Length,iris$Petal.Width)[1, 1]
slope = slr_beta(iris$Petal.Length,iris$Petal.Width)[2, 1]
abline(intercept, slope)
```
