---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document introduces the usage slrmodel’s basic set of functions and compares the slrmodel functions with the lm() function. To learn more about the lm() function, please refer to [lm() RDucumentation page](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm).

```{r setup}
library(slrmodel)
```

## Introduction
The simple linear regression model $y={\beta_0}+{\beta_1}x+{\epsilon}$ is a model that summarizes the relationship between a predictor variable `x` and a repsonse variable `y`. ${\beta_0}$ and ${\beta_1}$ are usually unknown and need to be estimated using observed data (x~1~,y~1~), (x~2~,y~2~),…, (x~n~,y~n~), and ${\epsilon}$ is a random error term. When the estimators are calculated using 

Deriving the estimators 
 and 
 by the method of least squares would help to obtain the best-fitting line 
, that is, the straight line demonstrating the best approximation of the given set of data.

The goal of slrmodel is to calculate the 
 and 
 using the method of least squares, test the significance of (x,y) association, and evaluate the overall fitness of the 
 model.

slrmodel aims to provide simple functions for each basic verb of data manipulation. These verbs can be organised into three categories based on the component of the dataset that they work with:

The simple linear regression model  is a model that summarizes the relationship between an independent variable x (predictor) a dependent variable y (response).  and  are usually unknown and need to be estimated using observed data (x1,y1), (x2,y2),…, (xn,yn), and  is a random error term. Deriving the estimators 
 and 
 by the method of least squares would help to obtain the best-fitting line 
, that is, the straight line demonstrating the best approximation of the given set of data.

The goal of slrmodel is to calculate the 
 and 
 using the method of least squares, test the significance of (x,y) association, and evaluate the overall fitness of the 
 model.

* `slr_beta()`
* `slr_rsquared()`

## Usage

This section is a detailed example of how to apply slrmodel in practices.

### Data: iris
To explore the functions of slrmodel, we will use the dataset `iris`. This dataset contains 150 cases (rows) and 5 variables (columns). More information about the dataset can be found in `?iris`.
```{r}
dim(iris)
head(iris)
```
Before continuing to analyze the data you must:

* Determine your target predictor variable and response variable.

* Extract the data of interest from the full dataset.

* Store the set of predictor variables and the set of response variables in two separate vectors.

### Create a matrix of simple linear regression coefficients with `slr_beta()`

`slr_beta()` allows you to select a subset of rows in a data frame. Like all single verbs, the first argument is the tibble (or data frame). The second and subsequent arguments refer to variables within that data frame, selecting rows where the expression is TRUE.
```{r}

```

### Calculate the coefficient of determination with `slr_rsquared()`
`slr_rsquared()` allows you to select a subset of rows in a data frame. Like all single verbs, the first argument is the tibble (or data frame). The second and subsequent arguments refer to variables within that data frame, selecting rows where the expression is TRUE.
```{r}

```

## A comparison between functions of slrmodel and the `lm()` function

```{r}
if(!require(bench)) install.packages("bench",repos = "http://cran.us.r-project.org")
bench::mark(
  slr_rsquared_result=slr_rsquared(iris$Petal.Width,iris$Petal.Length),
  lm_result=summary(lm(Petal.Length~Petal.Width,data=iris))$r.squared
)
```

